Landform boundary effects on Holocene forager landscape use in arid South Australia

### 2017-06-05 analysis/paper/paper.Rmd
do ratio of observed site distance to max-min possible dist

library(knitr)
opts_chunk$set(echo = FALSE,
               warning = FALSE,
               message = FALSE,
               dpi = 900)
options(warn = -1)

@@

```{r distanceofsitetoboundary, fig.cap="Distribution of distances of archaeological sites to terrain pattern bounaries. Distance is shows the vertical axis with a logarithmic scale.  Data points show distance values for individual sites.", fig.width=fig_width_one_col}

@@

# plot
library(ggplot2)
library(ggforce)
ggplot(ODX_allsites_2010.s, 
       aes(reorder(GREGLANDF, -RASTERVALU),
           RASTERVALU)) + 
  geom_boxplot() + 
  scale_y_log10(breaks = seq(0,1400,200)) + 
  theme_bw(base_size = text_size-1) +
  theme(axis.title.y = element_text(angle = 90))  +
  geom_sina(alpha = 0.03) +
  xlab("Terrain pattern") +
  ylab("Distance of site to boundary (m)")
```

>>>>

library(knitr)
opts_chunk$set(echo = FALSE,
               warning = FALSE,
               message = FALSE,
               dpi = 900,
               cache = TRUE)
options(warn = -1)

@@

```{r}
# get shapefiles
library(units)
library(sf)
odx_polygons <- read_sf("../data/ODX_TP_2010_data/CLipped_ODO_terrain_patterns_2010_Ja.shp")
```

```{r tpelongationandlength,  fig.cap="Upper: Distribution of elongation ratios of terrain pattern polygons in the study area. The ANOVA output in the upper right of the plot indicates that there is no significant difference in the elongation ratios of the terrain patterns. Lower: Distribution of calliper lengths (longest distance between two points) of the terrain patterns in the study area. The Krukall-Wallis test in the upper right of the plot indicates that there is no significant difference in the lengths of the terrain patterns.", fig.width=fig_width_two_cols}

# compute elongation of each landform area

# subset to get only TPs 1 km or more accross

odx_polygons <- 
  filter(odx_polygons,  GREGLANDF %in% c("A4", "Czs4", "Q4", "K4", "Czs5", "Q5", "Q6"))

# simplify to reduce slivers

library(rmapshaper)

odx_polygons_simpl <- vector("list", length = nrow(odx_polygons))
for(i in seq_len(nrow(odx_polygons))) {
  t1 <- ms_simplify(as(odx_polygons[i, ], 'Spatial'), keep = 0.2)
  t2 <- st_as_sf(t1)
  odx_polygons_simpl[[i]] <- t2
  # print(i)
}

# get coords of polygons to put into Momocs

odx_polygons_coords <- apply(odx_polygons, 1,  function(i) st_coordinates(i$geometry))

odx_polygons_coords <- lapply(odx_polygons_simpl, function(i) st_coordinates(i$geometry))

# create Momocs outlines object

library(Momocs)

odx_polygons_outline <- Out(odx_polygons_coords,
                            fac = data.frame(TP_NO = odx_polygons$TP_NO))

# compute elongation for each shape

odx_polygons_elongation <- 
  data.frame(GEOLREGIME = odx_polygons$GEOLREGIME,
             LF = odx_polygons$LANDFORM,
             GREGLANDF = odx_polygons$GREGLANDF,
             ID = odx_polygons$TP_NO,
             elongation = sapply(odx_polygons_outline$coo, coo_elongation),
             calliper_length = sapply(odx_polygons_outline$coo, coo_calliper))

# how varied are the shapes of each TP?

#  hist(odx_polygons_elongation$elong) it's normal
tp_elong_anova <- summary(aov(elongation ~ GREGLANDF, data = odx_polygons_elongation))
# no signifant difference

fval <- round(tp_elong_anova[[1]]$`F value`[1], 3)
pval <- round(tp_elong_anova[[1]]$`Pr(>F)`[1], 3)
df <-  paste0(tp_elong_anova[[1]]$Df, collapse = ", ")

anova_string <- paste0("F(", df, ") = ", fval, ", p = ", pval)

library(ggplot2)
elong <- ggplot(odx_polygons_elongation,
       aes(reorder(GREGLANDF, -elongation),
           elongation)) +
  geom_boxplot() +
  theme_bw(base_size = text_size-1) +
  xlab("") +
  ylab("Elongation ratio") +
  annotate("text",  x = "Q5", y = 0.85, label = anova_string)

# similar analysis for maximum dimension (calliper length or Feret's diameter)

#  hist(odx_polygons_elongation$calliper_length) it's NOT normal
tp_length_kruskal <- kruskal.test(calliper_length ~ GREGLANDF, data = odx_polygons_elongation)
# no signifant difference

hval <- unname(round(tp_length_kruskal$statistic, 3))
pval <- round(tp_length_kruskal$p.value, 3)
df <-  unname(tp_length_kruskal$parameter)

kruskal_string <- paste0("H(", df, ") = ", hval, ", p = ", pval)

library(ggplot2)
calliper <- ggplot(odx_polygons_elongation,
       aes(reorder(GREGLANDF, -calliper_length),
           calliper_length/1000)) +
  geom_boxplot() +
  theme_bw(base_size = text_size-1) +
  xlab("Terrain Pattern") +
  ylab("Calliper length (km)") +
  annotate("text",  x = "Q4", y = 18, label = kruskal_string)

library(gridExtra)
grid.arrange(elong, calliper, ncol = 1)

```{r distanceofsitetoboundary}

@@

# plot
library(ggplot2)
library(ggforce)
d2b <- ggplot(ODX_allsites_2010.s, 
       aes(reorder(GREGLANDF, -RASTERVALU),
           RASTERVALU)) + 
  geom_boxplot() + 
  scale_y_log10(breaks = seq(0,1400,200)) + 
  theme_bw(base_size = text_size-1) +
  theme(axis.title.y = element_text(angle = 90))  +
  geom_sina(alpha = 0.03) +
  xlab("Terrain pattern") +
  ylab("Distance of site to boundary (m)")
```

```{r maxmindistancetoboundary, fig.cap="Left: Distributions of distances of archaeological sites to terrain pattern bounaries. Right: Distributions of ratios of observed distance of site to terrain pattern boundary and the maximum possible distance to boundary for each terrain pattern. Distance is shown on the vertical axis with a logarithmic scale. Data points show distance values for individual sites.", fig.width=fig_width_two_cols}

# "a new variable should be created for each site which is a ratio of 
# minimum distance to boundary [of that site] to maximum possible 
# minimum distance to boundary for any location within a polygon."

library(dplyr)
library(tidyr)

# make a fn to ut random points on a polygon, and compute areas to boundary for all points in that polygon

library(sp)
library(rgeos)

random_points_to_boundary <- function(poly, n = 1000, ...){
  # convert st to sp poly <- odx_polygons[2,]
  t1 <- as(poly, 'Spatial')
  # generate random points in that polygon
  t2 <- spsample(t1, n = n,type = "random")
  # compute distance of points to polygon boundary
  gd <- gDistance(t2, as(t1, "SpatialLines"), byid = TRUE) 
  return(as.vector(unname(gd)))
}

# run that fn for each poly in the study area
odx_polygons_dists <- vector("list", length = nrow(odx_polygons))
for(i in seq_len(length(odx_polygons_dists))){
  odx_polygons_dists[[i]] <-random_points_to_boundary(odx_polygons[i,])
  # print(i)
}

names(odx_polygons_dists) <- odx_polygons$TP_NO

odx_polygons_dists <- as.data.frame(do.call('cbind', odx_polygons_dists))

odx_polygons_dists <- stack(odx_polygons_dists)

odx_polygons_dists <- separate(odx_polygons_dists, 
                               ind, c("TP", "LF"), 
                               sep = "/",
                               remove = FALSE)

odx_polygons_dists_max <- 
odx_polygons_dists %>% 
  group_by(TP) %>% 
  dplyr::summarise(values = max(values))

# ratio of actual site distance to boundary and max TP distance to boundary

odx_polygons_dists_LF_max <- 
  odx_polygons_dists %>% 
  group_by(ind) %>% 
  dplyr::summarise(values = max(values)) %>% 
  mutate(TP_NO = ind)

ODX_allsites_2010.s_with_rand <- 
  ODX_allsites_2010.s %>% 
    left_join(odx_polygons_dists_LF_max) %>% 
    mutate(ratio = RASTERVALU/values)

library(ggplot2)
library(ggforce)
r2b <- ggplot(ODX_allsites_2010.s_with_rand, 
       aes(reorder(GREGLANDF, -ratio),
           ratio)) + 
  geom_boxplot() + 
  scale_y_log10() + 
  theme_bw(base_size = text_size-1) +
  theme(axis.title.y = element_text(angle = 90))  +
  geom_sina(alpha = 0.03) +
  xlab("Terrain pattern") +
  ylab("Ratio of site distance to boundary \nto maximum possible distance to boundary")

grid.arrange(d2b, r2b, ncol = 2)
